Board Game Recommendation

### 2015-03-19 paper.Rnw
couple of suggestions, read-through section 3.2

\documentclass[twoside]{article}
\usepackage{multicol}
\usepackage[hmarginratio=1:1, top=32mm, columnsep=20pt]{geometry}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{url}

@@

\begin{document}
\maketitle

@@

There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section). The characteristics we're interested in fall into these categories as outlined in table~%\ref{tab:categories}

\hh{Just to make this flow better - make the table below an actual table in the table environment. Give it a caption and refer to it in the text ... I've just implemented it for you}
\begin{table}[ht]
\centering
\begin{tabular}{|l|c|}
\hline
Category &  This covers the broad categories a game\\
& might fall into (ex.: Humor, Puzzle, Sports, etc.) \\
\hline
Family &  Some games a naturally part of a "family" \\
&  that share a name or common element \\
& (ex.: Animals:Bats, Ancient Wars Series, Hello Kitty)\\
\hline
Mechanic &  This is what mechanisms the game uses\\
& (ex.: Dice Rolling, Card Drafting, Worker Placement)\\
\hline
Subdomain & More general than Category, but loosely bins games\\
& (ex.: Family Games, Strategy Games, etc.)\\
\hline
\end{tabular}
\caption{\label{tab:categories} Overview of the  relevant characteristics of board games considered in this project.}
\end{table}

<<classifiers, echo = FALSE, results = 'asis'>>=
xtable(paper_stuff$classifiers, caption = c("Classifiers Data, as seen in Database", "Classifiers Table"))
@

@@

Not all of the data was scraped before creating the applicaiton. Data for 6,088 games was scraped and stored across the four tables. The reason for not scraping all of the data is twofold. First of all, for those 6,088 games there are over 3 million unique ratings. If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippilingly large for the small desktop I have to work on. Secondly, this will force us to implement a continuous data scraping procedure. Each time the person uses the app, all we have to do is make sure we have all of their current information correct. We look through our ratings and can check them with their userpage on BoardGameGeek. From this, we can revise old ratings or add new ones, as well as check to see that all of the games that they've rated are in our database!

\section{Model}
\subsection{Model Selection}
To begin implementing such a model, we can look to two papers on one of the most famous recommender systems of all: Netflix. In 2006, Netflix offered a \$1 Million prize for a improvement on their algorithm for predicting user rating on movies they had not seen. In 2008, two teams: BigChaos and BellKor (the eventual winner), published papers on their methodology. In the end, they ended up taking something of a shotgun approach, using a linear blend of many different models to come up with a single super-model.

There are a few things that make sense to model first. Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings. These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games. We will implement these in a similar manner to the way that BigChaos implemented their Global Effects model \citep{bigchaos}. While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others. From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%. Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.

@@

Where for the first step, $r_{u,i}$ will be the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise. In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.

To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}). The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$. Either way, the structure is the same.
\begin{equation}
\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
\end{equation}

Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closes interval endpoint.
%  The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
% 
% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
% 
% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
% 
% \begin{center}
% \begin{align}
% y_{u,i} &= \mu_{global} + error_{global} \\
% error_{global} &= \mu_{i} + error_{game} \\
% error_{game} &= \mu_{u} + error_{user} 
% \end{align}
% \end{center}
% 
% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
% 
% \begin{center}
% \begin{align}
% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
% \end{align}
% \end{center}
% \hh{something went wrong in the formula - what are you summing over?}

@@

\subsubsection{Optimizing $\alpha$}
To optimize the $\alpha$ values, we'll take all of the ratings for all players and subtract off the mean of all of the ratings (this is the first step in the globalized residual process). This leaves us with three variables: The userid, the gameid and the residual. We'll randomly split all of the data into 10 equally sized buckets. We'll use each of these buckets as a test group, where we put the other buckets together, fit a model using a canidate value for $\alpha$, and then fit the values in the test group. From each test group, we'll then find the mean square error, and then average those together from all 10 groups. Taking the square root of that average, we can get the RMSE for each canidate value for $\alpha$. Finally, we'll proceed using the canidate value for $\alpha$ with the lowest RMSE. A graph of this process can be seen below, with the domain trunacted for easier visibility.

<<game_alpha_optim, echo = F>>=
load("game_graph")
game_graph
@

>>>>

\documentclass[twoside]{article}
\usepackage{multicol}
\usepackage[hmarginratio=1:1, top=32mm, columnsep=20pt]{geometry}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{url}
\usepackage[normalem]{ulem}

@@

\begin{document}
\maketitle
\begin{abstract}
\hh{You'll need to include an abstract here}
\end{abstract}

@@

There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section). The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}.

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|}
\hline
Category &  This covers the broad categories a game\\
& might fall into (ex.: Humor, Puzzle, Sports, etc.) \\
\hline
Family &  Some games a naturally part of a "family" \\
&  that share a name or common element \\
& (ex.: Animals:Bats, Ancient Wars Series, Hello Kitty)\\
\hline
Mechanic &  This is what mechanisms the game uses\\
& (ex.: Dice Rolling, Card Drafting, Worker Placement)\\
\hline
Subdomain & More general than Category, but loosely bins games\\
& (ex.: Family Games, Strategy Games, etc.)\\
\hline
\end{tabular}
\caption{\label{tab:categories} Overview of the  relevant characteristics of board games considered in this project.}
\end{table}

\hh{Refer to table~\ref{tab:classifiers} somewhere in the text, add a label to the ratings table as well.}
<<classifiers, echo = FALSE, results = 'asis'>>=
print(xtable(paper_stuff$classifiers, caption = c("Example lines of classifiers data, as seen in Database", "Classifiers Table"), label="tab:classifiers"), include.rownames =FALSE )
@

@@

Not all of the data was scraped before creating the applicaiton. Data for 6,088 games was scraped and stored across the four tables. The reason for not scraping all of the data is twofold. First of all, for those 6,088 games there are over 3 million unique ratings. If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippingly large for all but very high-powered machines. %the small desktop I have to work on. 
\hh{Instead, we} %Secondly, this will force us to 
implement a continuous data scraping procedure \hh{and update the recommender model dynamically:} Each time a person uses the app, %all we have to do is 
we make sure to check that all of their current information is correct. We look through known ratings and  scrape  their userpage on BoardGameGeek. Based on this, we  revise old ratings and add new ones, as well as check to see that all of the games that this user has rated are included in the database!

\section{Model}
\subsection{Model Selection}
\hh{We base our implementation of a recommender system} %To begin implementing such a model, we can look to 
on two papers  of on one the most famous recommender systems of all: Netflix. In 2006, Netflix offered a \$1 Million prize for an improvement on their algorithm for predicting user rating on movies they had not seen. In 2008, two teams: BigChaos and BellKor (the eventual winner), published papers on their methodology. In the end, these teams ended up taking something of a shotgun approach, using a linear blend of many different models to come up with a single super-model.

\hh{XXX What is the approach you are taking compared to Netflix? the same? similar? what is different? - One or two sentences here.}

There are a few things that make sense to model first. Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings. These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games. We will implement these in a similar manner to the way that BigChaos implemented the Global Effects model \citep{bigchaos}.
\hh{XXX the next clause needs an end to be a sentence :)}
While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others. 
\hh{XXX explain RMSE; make sure in the next paragraph to always point out what is yours and what is netflix. You know all those distinctions, but for any other reader it is not that obvious.}
From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%. Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.

@@

\hh{XXX what do you mean by first step? If you need iterations, use $r_{u,i}^{(n)}$ or something similar. If there is an iteration, make sure to give a convergence criterion.

Ah, ok - I see - I think what is going on is that you first get rid of the individuals' effects, and then of the games' effect. Write that out in two separate pieces. Give them appropriate names :)  - then you will have a much easier time describing what you are doing than in the paragraph below, where you are switching out us and is.
}
\hh{XXX: doesn't this error need to be indexed as well? could we call the error $\epsilon$ and make away with the text in the formula? }
\hh{XXX: generally:  no future tense in the write up. The CC is in the here and now :)
}

Where for the first step, $r_{u,i}$ \sout{will be} \hh{is} the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise. In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.

\hh{XXX it is $\theta_u$, not $\theta$. }
\hh{XXX Why do you call $\alpha$ a fitting parameter and not a shrinking parameter?}

To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shrunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}). The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$. Either way, the structure is the same. \hh{Again, write this out in two separate formulae.}
\begin{equation}
\widehat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
\end{equation}

Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closest interval endpoint.
%  The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
% 
% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
% 
% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
% 
% \begin{center}
% \begin{align}
% y_{u,i} &= \mu_{global} + error_{global} \\
% error_{global} &= \mu_{i} + error_{game} \\
% error_{game} &= \mu_{u} + error_{user} 
% \end{align}
% \end{center}
% 
% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
% 
% \begin{center}
% \begin{align}
% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
% \end{align}
% \end{center}
% \hh{something went wrong in the formula - what are you summing over?}

@@

\subsubsection{Optimizing $\alpha$}
To optimize the $\alpha$ values, we'll take all of the ratings for all players and subtract off the mean of all of the ratings (this is the first step in the globalized residual process). This leaves us with three variables: The userid, the gameid and the residual. We'll randomly split all of the data into 10 equally sized buckets. We'll use each of these buckets as a test group, where we put the other buckets together, fit a model using a canidate value for $\alpha$, and then fit the values in the test group. From each test group, we'll then find the mean square error, and then average those together from all 10 groups. Taking the square root of that average, we can get the RMSE for each canidate value for $\alpha$. Finally, we'll proceed using the canidate value for $\alpha$ with the lowest RMSE. A graph of this process can be seen \hh{in figure~\ref{fig:alpha}}, with the domain truncated for easier visibility.

\begin{figure}
<<game_alpha_optim, echo = F>>=
load("game_graph")
game_graph
@
\caption{\label{fig:alpha} You need a caption here - what is shown?}
\end{figure}
